The Proofs

### 2024-07-31 proofs/tex/translation-proofs.tex
fix \cref

      The premise of the given judgment is 
      $[[fΓ; fΦ, x:fA  ⊢ ft : fB  ⤳ c]]$, which by the induction hypothesis
      implies $[[+|fΓ|+ ; +|fΦ, x:fA|+  ⊢ c : ↑+|fB|+]]$,
      which we rewrite as 
      $[[+|fΓ|+ ; +|fΦ|+, x:+|fA|+  ⊢ c : ↑+|fB|+]]$.
      Also, the soundness of \systemf typing (together with \cref{lemma:pol-removable}) 
      imply that the inferred type $[[fA → fB]]$ is well-formed in $[[fΓ]]$,
      which means that $[[+|fA|+]]$ is well-formed in 
      $[[+|fΓ|+]]$ by \cref{lemma:typ-pol-wf-completeness}.
      Then by \ruleref{\ottdruleDTtLamLabel} we infer 
      $[[+|fΓ|+ ; +|fΦ|+  ⊢ λx : +|fA|+. c : +|fA|+ → ↑+|fB|+]]$,
      and then by \ruleref{\ottdruleDTThunkLabel} and \ruleref{\ottdruleDTReturnLabel}, 
      $[[+|fΓ|+ ; +|fΦ|+  ⊢ return {λx : +|fA|+. c} : ↑↓(+|fA|+ → ↑+|fB|+)]]$, 
      which can be rewritten to the required judgment since
      $[[↓(+|fA|+ → ↑+|fB|+) = +|fA → fB|+]]$ by definition.

@@

    \item \ruleref{\ottdruleELAppLabel}.
      By inversion of the given judgment, we have:
      \begin{itemize}
        \item $[[fΓ ; fΦ ⊢ ft : fA → fB  ⤳  c]]$,
          and by the induction hypothesis,
          $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑+|fA → fB|+]]$,
          which is equivalent to
          $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑↓(+|fA|+ → ↑+|fB|+)]]$.
        \item $[[fΓ ; fΦ ⊢ ft' : fA  ⤳  c']]$, 
          and by the induction hypothesis,
          $[[+|fΓ|+ ; +|fΦ|+ ⊢ c' : ↑+|fA|+]]$.
      \end{itemize}
      To infer 
      $[[+|fΓ|+ ; +|fΦ|+ ⊢ let f : +|fA → fB|+ = c; let x : +|fA|+ = c'; let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$, 
      we gradually apply the corresponding rules of \fexists to construct the required typing inference tree:
      \begin{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+, y:+|fB|+ ⊢ return y : ↑+|fB|+]]$ by 
          \ruleref{\ottdruleDTVarLabel} and \ruleref{\ottdruleDTReturnLabel}.  
        \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+ ⊢ let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ by 
          \ruleref{\ottdruleDTAppLetAnnLabel}. Let us show that the 
          required premises hold. 
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ +|fB|+]]$ by \cref{lemma:typ-pol-wf-completeness};
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+ ⊢ f : ↓(+|fA|+ → ↑+|fB|+)]]$ by 
              \ruleref{\ottdruleDTVarLabel};
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+ ⊢ (+|fA|+ → ↑+|fB|+) ● x ⇒> ↑+|fB|+]]$
              by \ruleref{\ottdruleDTArrowAppLabel};
            \item $[[+|fΓ|+ ⊢ ↑+|fB|+ ≤ ↑+|fB|+]]$ by reflexivity (\cref{lemma:subtyping-reflexivity});
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+, y:+|fB|+ ⊢ return y : ↑+|fB|+]]$ 
            as noted above;
          \end{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+) ⊢ let x:+|fA|+ = c'; let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ by 
          \ruleref{\ottdruleDTCVarLetLabel}. Let us show that the 
          required premises hold. 
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ +|fA|+]]$ by \cref{lemma:typ-pol-wf-completeness};
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+) ⊢ c' : ↑+|fA|+]]$
              holds by weakening of $[[+|fΓ|+ ; +|fΦ|+ ⊢ c' : ↑+|fA|+]]$, which in turn holds by the induction hypothesis,
              as noted above;
            \item $[[+|fΓ|+  ⊢ ↑+|fA|+ ≤ ↑+|fA|+]]$ by reflexivity (\cref{lemma:subtyping-reflexivity});
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+ ⊢ let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ as noted above;
          \end{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+ ⊢ let f : +|fA → fB|+ = c; let x : +|fA|+ = c'; let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ by 
          \ruleref{\ottdruleDTCVarLetLabel}. Let us show that the 
          required premises hold. 
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ +|fA → fB|+]]$ by \cref{lemma:typ-pol-wf-completeness};
            \item $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑+|fA → fB|+]]$ by the induction hypothesis, as noted above;
            \item $[[+|fΓ|+  ⊢ ↑+|fA → fB|+ ≤ ↑+|fA → fB|+]]$ by reflexivity (\cref{lemma:subtyping-reflexivity});
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+) ⊢ let x:+|fA|+ = c'; let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ 
              as noted above (we rewrote $[[+|fA → fB|+]]$ as $[[↓(+|fA|+ → ↑+|fB|+)]]$).
          \end{enumerate}
      \end{enumerate}

    \item \ruleref{\ottdruleELTAppLabel}
      By inversion of this rule, we have
      \begin{itemize}
        \item $[[fΓ ; fΦ ⊢ ft : ∀a.fT  ⤳  c]]$, and thus, 
          $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑+|∀a.fT|+]]$ by the induction hypothesis; and
        \item $[[fΓ ⊢ fA]]$, which implies $[[+|fΓ|+ ⊢ +|fA|+]]$ by \cref{lemma:typ-pol-wf-completeness}.
      \end{itemize}
      We wish to infer
      $[[+|fΓ|+ ; +|fΦ|+ ⊢ let f : +|∀a.fT|+ = c; let y : +|[fA/a]fT|+ = f(·); return y : ↑+|[fA/a]fT|+]]$.
      First, let us rewrite the polarized types by definition and \cref{lemma:pol-subst}:
      $[[+|fΓ|+ ; +|fΦ|+ ⊢ let f : ↓∀α⁺.↑+|fT|+ = c; let y : [+|fA|+/α⁺]+|fT|+ = f(·); return y : ↑[+|fA|+/α⁺]+|fT|+]]$.
      Then we prove this judgment gradually by applying the corresponding rules of \fexists:
      \begin{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+, y : [+|fA|+/α⁺]+|fT|+ ⊢ return y : ↑[+|fA|+/α⁺]+|fT|+]]$
          holds by \ruleref{\ottdruleDTVarLabel} and \ruleref{\ottdruleDTReturnLabel}.
        \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+  ⊢ let y : [+|fA|+/α⁺]+|fT|+ = f(·); return y : ↑[+|fA|+/α⁺]+|fT|+]]$
          holds by \ruleref{\ottdruleDTAppLetAnnLabel}. Let us show that the 
          required premises hold. 
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ [+|fA|+/α⁺]+|fT|+]]$ holds since $[[+|fΓ|+ ⊢ +|fA|+]]$ and $[[+|fΓ|+ ⊢ +|fT|+]]$ by \cref{lemma:wf-subst}.
            \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+  ⊢ f : ↓∀α⁺.↑+|fT|+]]$ by \ruleref{\ottdruleDTVarLabel}.
            \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+  ⊢ ∀α⁺.↑+|fT|+ ● · ⇒> ∀α⁺.↑+|fT|+]]$ by \ruleref{\ottdruleDTEmptyAppLabel}.
            \item $[[+|fΓ|+  ⊢ ∀α⁺.↑+|fT|+ ≤ ↑[+|fA|+/α⁺]+|fT|+ ]]$ by \ruleref{\ottdruleDOneForallLabel}:
              $[[+|fΓ|+  ⊢ +|fA|+/α⁺ : {α⁺}]]$ since $[[+|fΓ|+  ⊢ +|fA|+]]$, as noted above.
            \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+, y : [+|fA|+/α⁺]+|fT|+ ⊢ return y : ↑[+|fA|+/α⁺]+|fT|+]]$ as noted above.
          \end{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+  ⊢ let f : ↓∀α⁺.↑+|fT|+ = c; let y : [+|fA|+/α⁺]+|fT|+ = f(·); return y : ↑[+|fA|+/α⁺]+|fT|+]]$
          holds by \ruleref{\ottdruleDTCVarLetLabel}. Let us show that the
          required premises hold.
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ ↓∀α⁺.↑+|fT|+]]$ by \cref{lemma:typ-pol-wf-completeness}.
            \item $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑↓∀α⁺.↑+|fT|+]]$ by the induction hypothesis, as noted above.
            \item $[[+|fΓ|+  ⊢ ↑↓∀α⁺.↑+|fT|+ ≤ ↑↓∀α⁺.↑+|fT|+]]$ by reflexivity (\cref{lemma:subtyping-reflexivity}).
            \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+  ⊢ let y : [+|fA|+/α⁺]+|fT|+ = f(·); return y : ↑[+|fA|+/α⁺]+|fT|+]]$ as noted above.
          \end{enumerate}
      \end{enumerate}
  \end{caseof}
\end{proof}

>>>>

      The premise of the given judgment is 
      $[[fΓ; fΦ, x:fA  ⊢ ft : fB  ⤳ c]]$, which by the induction hypothesis
      implies $[[+|fΓ|+ ; +|fΦ, x:fA|+  ⊢ c : ↑+|fB|+]]$,
      which we rewrite as 
      $[[+|fΓ|+ ; +|fΦ|+, x:+|fA|+  ⊢ c : ↑+|fB|+]]$.
      Also, the soundness of \systemf typing (together with \cref{lemma:pol-removable}) 
      imply that the inferred type $[[fA → fB]]$ is well-formed in $[[fΓ]]$,
      which means that $[[+|fA|+]]$ is well-formed in 
      $[[+|fΓ|+]]$ by \cref{lemma:typ-pol-wf}.
      Then by \ruleref{\ottdruleDTtLamLabel} we infer 
      $[[+|fΓ|+ ; +|fΦ|+  ⊢ λx : +|fA|+. c : +|fA|+ → ↑+|fB|+]]$,
      and then by \ruleref{\ottdruleDTThunkLabel} and \ruleref{\ottdruleDTReturnLabel}, 
      $[[+|fΓ|+ ; +|fΦ|+  ⊢ return {λx : +|fA|+. c} : ↑↓(+|fA|+ → ↑+|fB|+)]]$, 
      which can be rewritten to the required judgment since
      $[[↓(+|fA|+ → ↑+|fB|+) = +|fA → fB|+]]$ by definition.

@@

    \item \ruleref{\ottdruleELAppLabel}.
      By inversion of the given judgment, we have:
      \begin{itemize}
        \item $[[fΓ ; fΦ ⊢ ft : fA → fB  ⤳  c]]$,
          and by the induction hypothesis,
          $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑+|fA → fB|+]]$,
          which is equivalent to
          $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑↓(+|fA|+ → ↑+|fB|+)]]$.
        \item $[[fΓ ; fΦ ⊢ ft' : fA  ⤳  c']]$, 
          and by the induction hypothesis,
          $[[+|fΓ|+ ; +|fΦ|+ ⊢ c' : ↑+|fA|+]]$.
      \end{itemize}
      To infer 
      $[[+|fΓ|+ ; +|fΦ|+ ⊢ let f : +|fA → fB|+ = c; let x : +|fA|+ = c'; let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$, 
      we gradually apply the corresponding rules of \fexists to construct the required typing inference tree:
      \begin{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+, y:+|fB|+ ⊢ return y : ↑+|fB|+]]$ by 
          \ruleref{\ottdruleDTVarLabel} and \ruleref{\ottdruleDTReturnLabel}.  
        \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+ ⊢ let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ by 
          \ruleref{\ottdruleDTAppLetAnnLabel}. Let us show that the 
          required premises hold. 
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ +|fB|+]]$ by \cref{lemma:typ-pol-wf};
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+ ⊢ f : ↓(+|fA|+ → ↑+|fB|+)]]$ by 
              \ruleref{\ottdruleDTVarLabel};
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+ ⊢ (+|fA|+ → ↑+|fB|+) ● x ⇒> ↑+|fB|+]]$
              by \ruleref{\ottdruleDTArrowAppLabel};
            \item $[[+|fΓ|+ ⊢ ↑+|fB|+ ≤ ↑+|fB|+]]$ by reflexivity (\cref{lemma:subtyping-reflexivity});
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+, y:+|fB|+ ⊢ return y : ↑+|fB|+]]$ 
            as noted above;
          \end{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+) ⊢ let x:+|fA|+ = c'; let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ by 
          \ruleref{\ottdruleDTCVarLetLabel}. Let us show that the 
          required premises hold. 
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ +|fA|+]]$ by \cref{lemma:typ-pol-wf};
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+) ⊢ c' : ↑+|fA|+]]$
              holds by weakening of $[[+|fΓ|+ ; +|fΦ|+ ⊢ c' : ↑+|fA|+]]$, which in turn holds by the induction hypothesis,
              as noted above;
            \item $[[+|fΓ|+  ⊢ ↑+|fA|+ ≤ ↑+|fA|+]]$ by reflexivity (\cref{lemma:subtyping-reflexivity});
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+), x:+|fA|+ ⊢ let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ as noted above;
          \end{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+ ⊢ let f : +|fA → fB|+ = c; let x : +|fA|+ = c'; let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ by 
          \ruleref{\ottdruleDTCVarLetLabel}. Let us show that the 
          required premises hold. 
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ +|fA → fB|+]]$ by \cref{lemma:typ-pol-wf};
            \item $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑+|fA → fB|+]]$ by the induction hypothesis, as noted above;
            \item $[[+|fΓ|+  ⊢ ↑+|fA → fB|+ ≤ ↑+|fA → fB|+]]$ by reflexivity (\cref{lemma:subtyping-reflexivity});
            \item $[[+|fΓ|+ ; +|fΦ|+, f:↓(+|fA|+ → ↑+|fB|+) ⊢ let x:+|fA|+ = c'; let y:+|fB|+ = f(x); return y : ↑+|fB|+]]$ 
              as noted above (we rewrote $[[+|fA → fB|+]]$ as $[[↓(+|fA|+ → ↑+|fB|+)]]$).
          \end{enumerate}
      \end{enumerate}

    \item \ruleref{\ottdruleELTAppLabel}
      By inversion of this rule, we have
      \begin{itemize}
        \item $[[fΓ ; fΦ ⊢ ft : ∀a.fT  ⤳  c]]$, and thus, 
          $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑+|∀a.fT|+]]$ by the induction hypothesis; and
        \item $[[fΓ ⊢ fA]]$, which implies $[[+|fΓ|+ ⊢ +|fA|+]]$ by \cref{lemma:typ-pol-wf}.
      \end{itemize}
      We wish to infer
      $[[+|fΓ|+ ; +|fΦ|+ ⊢ let f : +|∀a.fT|+ = c; let y : +|[fA/a]fT|+ = f(·); return y : ↑+|[fA/a]fT|+]]$.
      First, let us rewrite the polarized types by definition and \cref{lemma:pol-subst}:
      $[[+|fΓ|+ ; +|fΦ|+ ⊢ let f : ↓∀α⁺.↑+|fT|+ = c; let y : [+|fA|+/α⁺]+|fT|+ = f(·); return y : ↑[+|fA|+/α⁺]+|fT|+]]$.
      Then we prove this judgment gradually by applying the corresponding rules of \fexists:
      \begin{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+, y : [+|fA|+/α⁺]+|fT|+ ⊢ return y : ↑[+|fA|+/α⁺]+|fT|+]]$
          holds by \ruleref{\ottdruleDTVarLabel} and \ruleref{\ottdruleDTReturnLabel}.
        \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+  ⊢ let y : [+|fA|+/α⁺]+|fT|+ = f(·); return y : ↑[+|fA|+/α⁺]+|fT|+]]$
          holds by \ruleref{\ottdruleDTAppLetAnnLabel}. Let us show that the 
          required premises hold. 
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ [+|fA|+/α⁺]+|fT|+]]$ holds since $[[+|fΓ|+ ⊢ +|fA|+]]$ and $[[+|fΓ|+ ⊢ +|fT|+]]$ by \cref{lemma:wf-subst}.
            \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+  ⊢ f : ↓∀α⁺.↑+|fT|+]]$ by \ruleref{\ottdruleDTVarLabel}.
            \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+  ⊢ ∀α⁺.↑+|fT|+ ● · ⇒> ∀α⁺.↑+|fT|+]]$ by \ruleref{\ottdruleDTEmptyAppLabel}.
            \item $[[+|fΓ|+  ⊢ ∀α⁺.↑+|fT|+ ≤ ↑[+|fA|+/α⁺]+|fT|+ ]]$ by \ruleref{\ottdruleDOneForallLabel}:
              $[[+|fΓ|+  ⊢ +|fA|+/α⁺ : {α⁺}]]$ since $[[+|fΓ|+  ⊢ +|fA|+]]$, as noted above.
            \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+, y : [+|fA|+/α⁺]+|fT|+ ⊢ return y : ↑[+|fA|+/α⁺]+|fT|+]]$ as noted above.
          \end{enumerate}
        \item $[[+|fΓ|+ ; +|fΦ|+  ⊢ let f : ↓∀α⁺.↑+|fT|+ = c; let y : [+|fA|+/α⁺]+|fT|+ = f(·); return y : ↑[+|fA|+/α⁺]+|fT|+]]$
          holds by \ruleref{\ottdruleDTCVarLetLabel}. Let us show that the
          required premises hold.
          \begin{enumerate}
            \item $[[+|fΓ|+ ⊢ ↓∀α⁺.↑+|fT|+]]$ by \cref{lemma:typ-pol-wf}.
            \item $[[+|fΓ|+ ; +|fΦ|+ ⊢ c : ↑↓∀α⁺.↑+|fT|+]]$ by the induction hypothesis, as noted above.
            \item $[[+|fΓ|+  ⊢ ↑↓∀α⁺.↑+|fT|+ ≤ ↑↓∀α⁺.↑+|fT|+]]$ by reflexivity (\cref{lemma:subtyping-reflexivity}).
            \item $[[+|fΓ|+ ; +|fΦ|+, f : ↓∀α⁺.↑+|fT|+  ⊢ let y : [+|fA|+/α⁺]+|fT|+ = f(·); return y : ↑[+|fA|+/α⁺]+|fT|+]]$ as noted above.
          \end{enumerate}
      \end{enumerate}
  \end{caseof}
\end{proof}
